\documentclass[a4 paper, 12pt]{article}
\usepackage{hyperref}
\usepackage{float}
\usepackage[pdftex]{graphicx}
\begin{document}
\author{Neil McCaffrey}
\title{A Look At The History Of The Gimp And Its Impact On Open Source
Software}
\date{\today}
\maketitle

%Ok so what I want to say in this report:
%how the gimp started and what motivated its inseption.
%show how a comunity got built around this peice of software
%give some insite as to the development of the gimp during the mid 2000s. What
%choices were made that influenced the success/failure of the project. What
%projects have come from the development of the gimp? Is the gimp still the
%flagship opensource project that it always was?

%Introduce the thing
	There have been few open source applications that have received as much
	attention in the media as the GIMP (GNU Image Manipulation Program). The
	community that has spawned around the GIMP is large and vibrant, with users
	sharing tutorials, scripts and plug-ins with each other. The project has been
	compared to large commercial software like Adobe Photoshop.  Although these
	comparisons have been unfavorable I want to make the argument that the GIMP
	is more important than just being a free image editor.

	The GIMP started life as a college assignment for students Spencer Kimball
	and Peter Mattis in 1995. Rather than write a compiler in Scheme/Lisp they
	decided to write an image manipulation program. They worked on the GIMP for
	about 10 months before releasing a working program in January of 1996. This
	initial release had features such as basic channel operation and a plug-in
	architecture which allowed users to extend the functionality of the
	program.

	However the first release was far from perfect. The dependency on the Motif
	toolkit for rendering GUI widgets made distribution of the GIMP
	difficult.The use of Motif also made the GIMP prone to crashing.  Because
	of these problems Peter got fed up with the Motif toolkit and started work
	on the GIMP Tool Kit or GTK for the 0.6x series of the GIMP.  GTK was then
	made object orientated with the release of GTK+.
	%Need to fill in the blank drawn here, too bored right now.

	In June 1997 Peter and Spencer left the project, having graduated and gotten
	jobs they no longer had time to contribute to the project. This left the
	project in a state of confusion. Patches were contributed but not merged
	into the official release. This lead to unofficial ``pre-releases'' to be
	made and published. To pick up the slack for this transition period Federico
	Mena Quintero took the lead of the project and lead it in the direction of
	stability and usability for the 1.0 release.

	With the founders of the project gone the development of the GIMP split up
	into teams developing the toolkit and a team developing the application.
	Where the project was to go next was discussed on #gimp. There was an
	obvious need for centralisation of both the code and documentation so Shawn
	Amundson secured the domain gimp.org. The developers started to manage their
	code using CVS around this time.

	%Feels like there is a bit of a gap here too. :/
	Despite its success in the open source world commercial users were unsatisfied
	by the GIMP. This lead to a fork of the 1.0.4 version called CinePaint in
	1999.\cite{release} CinePaint is an application which has very different
	goals to the GIMP. It has a much higher colour depth in comparison to the
	GIMP, 32bit vs 8bit per colour channel. With high colour depth and advanced
	features such as onion skinning CinePaint is an application focused on film
	retouching. It has seen use in many films such as `Spider Man', `Lord of the
	Rings' and `Harry Potter and the Philosopher's Stone'. It is second only to
	Photoshop in the film industry for retouching still images from films.
	
	

	As a result of its success the GIMP has often been compared to Adobe's
	Photoshop. This left many new users disappointed when they realised it is a
	very different application. In response forks of the GIMP have been made
	which attempt to replicate the Photoshop interface using the GIMP. GimpShop
	was first released in 2006 and was based on the 2.2.11 version of the GIMP.
	It featured menus which resembled those found in Photoshop and application
	windows were contained within a larger window like Photoshop. However
	development of the GimpShop stagnated soon after its release.

	Although having feature parity with Photoshop is not one of the GIMP
	developers' goals users have desired the application to have support for
	higher colour bit depth and CYMK colour channels. Rather than build this
	functionality directly into the application the developers decided to build
	an external library to handle these functions. This is the GEneric Graphics
	Library.(GEGL)
\begin{thebibliography}{12}
	\bibitem{earlyHistory}
		\url{``http://www.gimp.org/about/ancient_history.html''}
	\bibitem{gnomeHistory}
		\url{``http://primates.ximian.com/~miguel/gnome-history.html''}
	\bibitem{release}
		\url{``http://www.gimp.org/about/history.html''}
\end{thebibliography}
\end{document}
